### 2010-07-13 ms/main.tex
changed mass initiation, litte change in dat2punto.sh

This work aimes at the numerical simluation of high enthalpy flow in
porosities. The subject is of interest, as CITATION
NEEDED (Prof Hornung) has discovered porous surfaces to inhibit/block a
certain form of instabilities in hypersonic flows over a body/cone. A numerical
simulation of this configuration is desired to furnish data for a detailed
analysis of the acting phenomena. The simulation method of choice for the first approach
to this configuration is called Smoothed Particle Hydrodynamics (SPH).Although
there has been no known application of SPH to this precise/concrete case yet/before,
SPH is adopted, as it promises a crucial advantage in the way boundary
conditions are implemented. And for the treatment of porosities,
boundaries/walls represent a non negligible part of the problem.

SPH designates a meshfree method for numerical simulation of particle
ensembles (respectively matter that can be modelled as an ensemble of
particles as it is the case for fluids).
Thereby,the term particle ensembles not only refers
to fluids (as the name of the method would suggest) but also to solids. In
fact nowadays SPH is employed to simulate problems in different fields,
such as body elasticity and fracture,liquid flows and gas
dynamics~\cite{Monaghan2005}.
The initial developement of the method dates back to 1977, where Gingold and
Monaghan tempted to compute an astrophysical problem with a new, more
efficient method~\cite{Gingold1977}. Independently and simultaneously, work in
this field was also carried out by Lucy who came up with the same
idea~\cite{Lucy1977}.
ADVANTAGES???DISADVANTAGES???

@@

INTRODUCE THE TWO CODES AND MENTION THAT THEY ARE WRITTEN IN C/C++...

@@

To illustrate the basic idea of SPH, it is helpful to first have a look at the
ruling equations of fluid dynamics, here represented by the SPH equations in
lagrangian form. The use of the lagrangian form seems logical as SPH is a
particle method.

@@

\begin {equation}
\label{eq:EFD_form}
{\frac{dA(r)}{dt}}=f(A(r),\nabla A(r),r)
\end {equation}
where
\begin {equation}
{\frac{d(.)}{dt}}={\frac{\partial(.)}{ \partial t}}+v\nabla(.)
\end{equation}

@@

Applying the particle approximation on \ref{eq:int_interpolant} leads with the
relation between mass, density and volume for each discretized particle
\begin{equation}
m_i=\rho_i V_i
\end{equation}
 to the following expression

@@

In the same way, the derivative of the fieldfunction $frac{dA(x)}{dx}$ can be
approximated to give\cite{Monaghan2005}or\cite{Liu2003}(in Liu with gradient,
im Monaghan for 1 direction)

@@

The disadvantage of the above equation \ref{eq:simpleDerivative} is, that for
$A(r)=const.$ the approximated derivative does not vanish. By writing (in
Monaghan the equation figures not with gradient but only in 1 direction!?!)
\begin{equation}
\label{eq:derivativeIdentity}
\nabla A = \frac{1}{\Phi}(\nabla {\Phi A}-A\nabla \Phi)
\end{equation}
where $\Phi$ is any differentiable function, this problem can be resolved. In its
SPH form  equation \ref{eq:derivativeIdentity} reads as follows:
\begin{equation}
\nabla_a A = \frac{1}{\Phi_a}\sum_b \frac{m_b}{\rho_b}(A_b-A_a)\nabla_a W_{ab}
\end{equation}

It can be seen, that the (approximation of the) derivative of the
field function $A$ is now expressed by a summation over the fieldfunction itself
and the derivative of the kernel function - which can be determined exactly.

Now the basic formulations are introduced in principle: Both the fieldfunction
itself and its spacial derivatives can be approximated using only information
(=the states) of a discrete number of particles. Hence, the change rates of
the states described by the equations of fluid dynamics, which have a form
according to \ref{eq:EFD_form} can be rewritten using this information. Doing
this one obtains the SPH formulation of the equations of fluid dynamics.

The detailed derivation of those equations will not be shown here, as the
interest  only lies in describing the principle ideas behind the SPH
method. To resume, this are the following ones: A (field) function can be
described as an integral interpolation over the function itself and a kernel
function (which has to be chosen in an appropriate way). A further
approximation consists in transforming the integral over the domain into a
summation over the discrete elements constituing this domain (the particles). This allows to express both
the field function and its spacial derivatives dependent exclusively on discrete values
of the fieldfunction itself (and the kernel function, which is known as well as
its derivatvies).By finally choosing an appropriate kernel function with
compact support, the summation over all particles in the domain can be reduced
to a summation over those particles within the support domain of the particle
in question.

As already mentionned above, the SPH formulation of the equations of fluid
dynamics is obtained by applying the above principles and conducting different
transformations to the original (exact) equations. Depending on the
transformations used, the final SPH equations can assume a different form. In
the following, a common form for the change rates of density, velocity and
internal energy based on the EULER EQUATIONS (i.e. without viscosity) is
listed\cite{Monaghan2005}\cite{Liu2003} (perhaps also list NSG-form, and
explain why Euler is of interest as well (for 1DSPH code, perhaps later real
code including viscosity???)

@@

mention somewhere that kernel unity is inverse of volume?!?!?

@@

ILLUSTRATION OF VELOCITY ASSIGNMENT FOR VIRTUAL BOUNDARY PARTICLES

@@

DIAGRAM WITH CELLS /PARTICLES /DOMAIN!?!

@@

...equation of state in the original form...

...
transformation steps (or at least realations needed)...

@@

As the size of the 1D SPH program is not too big, it admits of the introduction/description of every single function used. The textual description is not complete but only comprises particularities and features that are considered worth mentionning. A Nassi-Schneidermann-Diagram (NSD)~\cite{Nassi???} complements each description and shows the control flow for each function as well as the interdependence of the individual functions. For the sake of a clear illustration , the NSD does not display every loop of the program; frequently, loops over all particles and the interaction list are ommitted (the instruction 'calculate ... for each particle' for example means the implementation of a loop over all particles i.e. all components of the corresponding variable's vector).  Consulting the comments in the source code is recommended if an even more detailed understanding is necessary or desired.

\paragraph{Explanation of a Nassi-Schneidermann-Diagram}
A NSD Diagram is employed in this section as it allows for a more compact description of the code structure within limited space than a conventional flow diagram and, in addition, offers a range of further advantages~\cite{Nassi???}. It is read from top to bottom, with different box symbols representing different kinds of instructions or control elements. The symbols used during the description of the functions below are the following ones:

>>>>

This work aimes at the numerical simluation of high enthalpy flow in
porosities. The subject is of interest, as CITATION
NEEDED (Prof Hornung) has discovered porous surfaces to inhibit/block a
certain form of instabilities in hypersonic flows over a body/cone. A numerical
simulation of this configuration is desired to furnish data for a detailed
analysis of the acting phenomena. The simulation method of choice for the first approach
to this configuration is called Smoothed Particle Hydrodynamics (SPH). Although
there has been no known application of SPH to this precise/concrete case yet/before,
SPH is adopted, as it promises a crucial advantage in the way boundary
conditions are implemented. And for the treatment of porosities,
boundaries/walls represent a non negligible part of the problem.

SPH designates a meshfree method for numerical simulation of particle
ensembles (respectively matter that can be modelled as an ensemble of
particles as it is the case for fluids).
Thereby,the term particle ensembles not only refers
to fluids (as the name of the method would suggest) but also to solids. In
fact nowadays SPH is employed to simulate problems in different fields,
such as body elasticity and fracture,liquid flows and gas
dynamics~\cite{Monaghan2005}.
The initial developement of the method dates back to 1977, where Gingold and
Monaghan tempted to compute an astrophysical gas dynamics problem with a new, more
efficient method~\cite{Gingold1977}. Independently and simultaneously, work in
this field was also carried out by Lucy who came up with the same
idea~\cite{Lucy1977}.
ADVANTAGES???DISADVANTAGES???

@@

INTRODUCE THE TWO CODES AND MENTION THAT THEY ARE CODED IN C/C++...

@@

To illustrate the basic idea of SPH, it is helpful to first have a look at the
ruling equations of fluid dynamics, here represented by the Navier stokes equations in
lagrangian form (using the substantive derivative). The use of the lagrangian form seems logical as SPH is a
particle method.

@@

\begin {equation}
\label{eq:EFD_form}
{\frac{dA(r)}{dt}}=f(A(r),\nabla A(r),r)
\end {equation}
where
\begin {equation}
{\frac{d(.)}{dt}}={\frac{\partial(.)}{ \partial t}}+v\nabla(.)
\end{equation}
is the substantive derivative.

@@

With the relation between mass, density and volume for each discretized particle
\begin{equation}
m_i=\rho_i V_i
\end{equation},
applying the particle approximation on \ref{eq:int_interpolant} leads to the following expression

@@

In the same way, the derivative of the field function $frac{dA(x)}{dx}$ can be
approximated to give\cite{Monaghan2005}or\cite{Liu2003}(in Liu with gradient,
im Monaghan for 1 direction)

@@

The disadvantage of the above equation \ref{eq:simpleDerivative} is, that for
$A(r)=const.$ the approximated derivative does not vanish. By writing (in
Monaghan the equation figures not with gradient but only in 1 direction!?!)
\begin{equation}
\label{eq:derivativeIdentity}
\nabla A = \frac{1}{\Phi}(\nabla {\Phi A}-A\nabla \Phi)
\end{equation}
where $\Phi$ is any differentiable function, this problem can be resolved. In its
SPH form  equation \ref{eq:derivativeIdentity} reads as follows:

\begin{equation}
\nabla_a A = \frac{1}{\Phi_a}\sum_b \frac{m_b}{\rho_b}(A_b-A_a)\nabla_a W_{ab}
\end{equation}

It can be seen, that the (approximation of the) derivative of the
field function $A$ is now expressed by a summation over the field function itself
and the derivative of the kernel function - which can be determined exactly.

Now the basic formulations are introduced in principle: Both the fieldfunction
itself and its spacial derivatives can be approximated using only information of a discrete number of particles. Hence, the change rates of
the states described by the equations of fluid dynamics, which have a form
according to \ref{eq:EFD_form} can be rewritten using this information. Doing
this one obtains the SPH formulation of the equations of fluid dynamics.

The detailed derivation of those equations will not be shown here, as the
interest  only lies in describing the principle ideas behind the SPH
method. To resume, this are the following ones: A (field) function can be
described as an integral interpolation over the function itself and a kernel
function (which has to be chosen in an appropriate way). A further
approximation consists in transforming the integral over the domain into a
summation over the discrete elements constituing this domain (the particles). This allows to express both
the field function and its spacial derivatives dependent exclusively on discrete values
of the fieldfunction itself (and of the kernel function, which is known as well as
its derivatvies). By finally choosing an appropriate kernel function with
compact support, the summation over all particles in the domain can be reduced
to a summation over those particles within the support domain of the particle
in question.

As already mentionned above, the SPH formulation of the equations of fluid
dynamics is obtained by applying the above principles and conducting different
transformations to the original (exact) equations. Depending on the
transformations used, the final SPH equations can assume a different form. In
the following, a common form for the change rates of density, velocity and
internal energy based on the EULER EQUATIONS (i.e. without viscosity) is
listed\cite{Monaghan2005,Liu2003} (perhaps also list NSG-form, and
explain why Euler is of interest as well (for 1DSPH code, perhaps later real
code including viscosity???)

@@

...mention somewhere that kernel unity is inverse of volume?!?!?

@@

FIGURE: ILLUSTRATION OF VELOCITY ASSIGNMENT FOR VIRTUAL BOUNDARY PARTICLES

@@

FIGURE: DIAGRAM WITH CELLS /PARTICLES /DOMAIN!?!

@@

%...equation of state in the original form...

%...
%transformation steps (or at least realations needed)...

@@

As the size of the 1D SPH program is not too big, it admits of the introduction/description of every single function used. The textual description is not complete but only comprises particularities and features that are considered worth mentionning. A Nassi-Schneidermann-Diagram (NSD)~\cite{Nassi1973} complements each description and shows the control flow for each function as well as the interdependence of the individual functions. For the sake of a clear illustration , the NSD does not display every loop of the program; frequently, loops over all particles and the interaction list are ommitted (the instruction 'calculate ... for each particle' for example means the implementation of a loop over all particles i.e. all components of the corresponding variable's vector).  Consulting the comments in the source code is recommended if an even more detailed understanding is necessary or desired.

\paragraph{Explanation of a Nassi-Schneidermann-Diagram}
A NSD Diagram is employed in this section as it allows for a more compact description of the code structure within limited space than a conventional flow diagram and, in addition, offers a range of further advantages~\cite{Nassi1973}. It is read from top to bottom, with different box symbols representing different kinds of instructions or control elements. The symbols used during the description of the functions below are the following ones:

### 2010-07-13 ms/titelblatt.tex
changed mass initiation, litte change in dat2punto.sh

\vspace*{-2.8cm}
\begin{minipage}[c]{.30\textwidth}
  \begin{flushleft}
    \includegraphics[width=3cm,clip=]{LOGO/new-aerlogo.eps}%
  \end{flushleft}
\end{minipage}
\begin{minipage}[c]{.43\textwidth}
    { ~\\ Institute of Aerodynamics \\ Prof.~Dr.-Ing. N.~A.~Adams}%
\end{minipage}
\begin{minipage}[c]{.25\textwidth}
  \begin{flushright}
    \vspace*{1em}
    \includegraphics[width=3cm,clip=]{LOGO/TUMLogo_oZ_Vollfl_blau_RGB.eps}%
  \end{flushright}
\end{minipage}

@@

\vspace*{1.2cm}
%\large
\normalsize
\vfill
\begin{tabular}{ll}
Betreuer: &  Dipl.-Ing. Sergey Litvinov \\
\ & \ \\
Ausgabe: & 15.\ Juni 2010\\
Abgabe: & 15.\ Dezember 2010\\
\end{tabular}

\vspace*{1.6cm}
%\large
Institute of Aerodynamics Technische Universit\"{a}t M\"{u}nchen\\
Graduate Aerospace Laboratories California Institute of Technology
2010	
\end{center}

>>>>

\vspace*{-2.8cm}
\begin{minipage}[c]{.30\textwidth}
  \begin{flushleft}
    \includegraphics[width=3cm,clip=]{LOGO/caltech_logo.eps}%
  \end{flushleft}
\end{minipage}
\begin{minipage}[c]{.43\textwidth}
\vspace*{1em}
    {GALCIT (Caltech)\\Prof. Hans G. Hornung \\ \\ Institute of Aerodynamics (TUM) \\ Prof.~Dr.-Ing. N.~A.~Adams}%
\end{minipage}
\begin{minipage}[c]{.25\textwidth}
  \begin{flushright}
    \vspace*{1em}
    \includegraphics[width=3cm,clip=]{LOGO/TUMLogo_oZ_Vollfl_blau_RGB.eps}%
  \end{flushright}
\end{minipage}

@@

\vspace*{1.2cm}
%\large
\normalsize
\vfill
\begin{tabular}{ll}
Betreuer: & ~Prof. Hans G. Hornung (Caltech)\\
\ & \ Dipl.-Ing. Sergey Litvinov (TUM)\\
 \\
Ausgabe: & 15.\ Juni 2010\\
Abgabe: & 15.\ Dezember 2010\\
\end{tabular}

\vspace*{1.6cm}
%\large
Institute of Aerodynamics Technische Universit\"{a}t M\"{u}nchen\\
Graduate Aerospace Laboratories California Institute of Technology\\
2010	
\end{center}
